---
title: "Soil temperature"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

#load packages

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
#library(chunked)
library(multcompView)
```

#Set working directory

```{r}
#setwd("C:/Users/cheet/OneDrive - Wageningen University & Research/thesis_josias_brian_shared/Josias and Brian")
path <- paste0("C:/Users/",  Sys.getenv("USERNAME"),"/OneDrive - Wageningen University & Research/thesis_josias_brian_shared/Josias and Brian/")
```

#Read in data

```{r}
soil <- read_csv(paste0(path, "/data/processed/button/soil.csv"))
```

# Create new columns for water amount and irrigation frequency

```{r}
soil_new <- soil %>%
  mutate(w_amount = case_when(
    treatment %in% c("i1", "i2", "i3") ~ 1,
    treatment %in% c("i4", "i5", "i6") ~ 0.65,
    TRUE ~ NA_real_
  )) %>% 
  mutate(
  ir_freq = case_when(
    treatment %in% c("i1", "i4") ~ 1,
    treatment %in% c("i2", "i5") ~ 2,
    treatment %in% c("i3", "i6") ~ 3,
    TRUE ~ NA_real_
  ))
soil_new$w_amount <- as.factor(soil_new$w_amount)
soil_new$ir_freq <- as.factor(soil_new$ir_freq)
```

#Create dataframe containing the start dates of irrigation for frequencies 2 and 3
```{r}
#Used for indicating irrigation timings and facilitate analysis
freq_2 <- c("2022-12-16 12:00:00", "2023-01-01 12:00:00", "2023-01-21 12:00:00", "2023-02-10 12:00:00")
freq_3 <- c("2022-12-22 12:00:00", "2023-01-07 12:00:00", "2023-01-27 12:00:00", "2023-02-10 12:00:00")
freq_2_datetime <- as.POSIXct(freq_2, format="%Y-%m-%d %H:%M:%S")
freq_3_datetime <- as.POSIXct(freq_3, format="%Y-%m-%d %H:%M:%S")
ir_dates <- data.frame(freq_2=freq_2_datetime, freq_3=freq_3_datetime, point2=c(21,21,21,21), point3=c(20,20,20,20))
```


#Plot temperature over time for each plot

```{r}
ggplot(soil_new, aes(x=time, y=temp)) +
  geom_point() +
  facet_wrap(~ Plot)
```

#p09 and p12 were removed from the soil and have bad data after a certain point

#Remove these data points

```{r}
# Filter data for the desired plot to determine at which point sensor was taken out of the soil
soil_plot09 <- soil_new %>% filter(Plot == "p09")
#view(soil_plot09)

soil_plot12 <- soil_new %>% filter(Plot == "p12")
#view(soil_plot12)
soil_plot20 <- soil_new %>% filter(Plot == "p20")
soil_plot25 <- soil_new %>% filter(Plot == "p25")
soil_plot48 <- soil_new %>% filter(Plot == "p48")


#big jump in temperature on 12.02.23 between 4:30 and 5:30 for both plots

# Create new column with filtered temperature values ignoring the values recorded out of the soil
#soil_filtered <- soil_new %>%
 # mutate(temp = case_when(
  #  Plot %in% c("p09", "p12") & time > "2023-02-12 04:30:27" ~ NA_real_,
   # TRUE ~ temp
  #))
soil_filtered <- soil_new %>%
  filter(!(Plot %in% c("p09", "p12") & time > "2023-02-12 04:30:27"))
```

```{r}
ggplot(soil_plot09, aes(x = time, y = temp)) +
  geom_line()
ggplot(soil_plot20, aes(x = time, y = temp)) +
  geom_line()
ggplot(soil_plot25, aes(x = time, y = temp)) +
  geom_line()
ggplot(soil_plot48, aes(x = time, y = temp)) +
  geom_line()
```

#Visualise data

```{r}
# Create plot for the filtered data (Each plot, different colour per treatment)
ggplot(soil_filtered, aes(x = time, y = temp, color = treatment)) +
  geom_point() +
  facet_wrap(~ Plot)
  

```

```{r}
# Create plot for the filtered data (Grouped by treatment)
ggplot(soil_filtered, aes(x = time, y = temp, color = treatment)) +
  geom_point() +
  #geom_line(aes(group = treatment)) +    #Line graph (not really useful)
  facet_grid(~ treatment)
```


#Plot mean daily temperature over time grouped by treatment

```{r}
# Calculate mean daily temperature and group by day
soil_summary <- soil_filtered %>%
  mutate(date = as.Date(time)) %>%
  group_by(date, treatment) %>%
  summarize(mean_temp = mean(temp, na.rm = TRUE))

# Create plot
ggplot(soil_summary, aes(x = date, y = mean_temp, color = treatment)) +
  geom_line() +
  labs(x = "Date", y = "Mean temperature", color = "Treatment") +
  scale_x_date(date_labels = "%m/%d") #+
  theme_minimal()
```

##Plot min daily temperature over time grouped by treatment (Not sure why i2 suddenly goes up. probably didn't filter errorneous data correctly.)

```{r}
# Calculate minimum daily temperature and group by day
soil_summary <- soil_filtered %>%
  mutate(date = as.Date(time)) %>%
  group_by(date, w_amount, ir_freq, treatment) %>%
  summarize(min_temp = min(temp, na.rm = TRUE),
            max_temp = max(temp, na.rm = TRUE),
            mean_temp = mean(temp, na.rm = TRUE)
            )
```

```{r}
# filter data to only include treatment i1
soil_i1 <- soil_summary %>% filter(treatment == "i1")

# calculate reference temperature for each date
soil_ref_temp <- soil_i1 %>% 
  group_by(date) %>%
  summarize(ref_min = min_temp,
            ref_mean = mean_temp,
            ref_max = max_temp)

# join original data to reference temperatures
soil_deltaT <- soil_summary %>% 
  left_join(soil_ref_temp, by = "date") %>% 
  mutate(deltaT_min = min_temp - ref_min,
         deltaT_mean = mean_temp - ref_mean,
         deltaT_max = max_temp - ref_max)

```


```{r}
# Create plot
ggplot(soil_summary, aes(x = date, color = w_amount)) +
  geom_point(aes(y = mean_temp), data = soil_summary) +
  geom_point(aes(y = min_temp), data = soil_summary) +
  geom_point(aes(y = max_temp), data = soil_summary) +
  ggtitle("Daily average, maximum, and minimum soil temperature") +
  xlab("Date") +
  ylab("Temperature (°C)")  +
  theme_minimal()
```

```{r}
# Create plot
ggplot(soil_summary, aes(x = date, color = w_amount)) +
  geom_line(aes(y = mean_temp), data = soil_summary) +
  geom_line(aes(y = min_temp), data = soil_summary) +
  geom_line(aes(y = max_temp), data = soil_summary) +
  facet_wrap(~ ir_freq,, nrow = 3) +
  ggtitle("Daily average, maximum, and minimum soil temperature") +
  xlab("Date") +
  ylab("Temperature (°C)")  +
  theme_minimal()
```


```{r}
# Create plot (Mean temp difference)
ggplot(soil_deltaT, aes(x = date, color = w_amount)) +
  geom_line(aes(y = deltaT_mean), data = soil_deltaT) +
  #geom_line(aes(y = deltaT_min), data = soil_deltaT) +
  #geom_line(aes(y = deltaT_max), data = soil_deltaT) +
  facet_wrap(~ ir_freq,, nrow = 3) +
  xlab("Date") +
  ylab("Temperature relative to L1I1 (°C)")  +
  theme_minimal() +
  guides(color = FALSE)
```


```{r}
# Create plot (Min temp difference)
ggplot(soil_deltaT, aes(x = date, color = w_amount)) +
  geom_line(aes(y = deltaT_min), data = soil_deltaT) +
  facet_wrap(~ ir_freq,, nrow = 3) +
  xlab("Date") +
  ylab("Temperature relative to L1I1 (°C)")  +
  theme_minimal() +
  guides(color = FALSE)
```

```{r}
# Create plot (Max temp difference)
ggplot(soil_deltaT, aes(x = date, color = w_amount)) +
  geom_line(aes(y = deltaT_max), data = soil_deltaT) +
  facet_wrap(~ ir_freq,, nrow = 3) +
  xlab("Date") +
  ylab("Temperature relative to L1I1 (°C)")  +
  theme_minimal() +
  guides(color = FALSE)

```

```{r}
#ANOVA
# Perform ANOVA
deltaT_mean_anova <- aov(deltaT_mean ~ w_amount*ir_freq + w_amount + ir_freq, data = soil_deltaT)
deltaT_min_anova <- aov(deltaT_min ~ w_amount*ir_freq + w_amount + ir_freq, data = soil_deltaT)
deltaT_max_anova <- aov(deltaT_max ~ w_amount*ir_freq + w_amount + ir_freq, data = soil_deltaT)
```

```{r}
summary(deltaT_mean_anova)
plot(deltaT_mean_anova)
```
```{r}
TukeyHSD(deltaT_mean_anova)

```


```{r}
summary(deltaT_min_anova)
plot(deltaT_min_anova)
```

```{r}
summary(deltaT_max_anova)
plot(deltaT_max_anova)
```

#Anova on daily min,max,mean temp
```{r}
#ANOVA
# Perform ANOVA
mean_temp_anova <- aov(mean_temp ~ w_amount*ir_freq + w_amount + ir_freq, data = soil_summary)
min_temp_anova <- aov(min_temp ~ w_amount*ir_freq + w_amount + ir_freq, data = soil_summary)
max_temp_anova <- aov(max_temp ~ w_amount*ir_freq + w_amount + ir_freq, data = soil_summary)
```

```{r}
# Summarize ANOVA results
summary(mean_temp_anova)
plot(mean_temp_anova)
#Bimodal or light-tailed
```
```{r}
shapiro.test(residuals(mean_temp_anova))
leveneTest(residuals(mean_temp_anova) ~ interaction(w_amount, ir_freq), data = soil_summary)
```
```{r}
# Summarize ANOVA results
summary(min_temp_anova)
plot(min_temp_anova)
#Bimodal? possibly light-tailed
```
```{r}
shapiro.test(residuals(min_temp_anova))
leveneTest(residuals(min_temp_anova) ~ interaction(w_amount, ir_freq), data = soil_summary)
```
```{r}
# Summarize ANOVA results
summary(max_temp_anova)
plot(max_temp_anova)
#light-tailed
```
```{r}
shapiro.test(residuals(max_temp_anova))
leveneTest(residuals(max_temp_anova) ~ interaction(w_amount, ir_freq), data = soil_summary)
```
#Non-parametric test
```{r}
df_daily_median <- soil_filtered %>%
  mutate(date = as.Date(time)) %>%
  group_by(treatment, date, w_amount, ir_freq) %>%
  summarize(median_temp = median(temp),
            max_temp = median(temp[temp == max(temp)]),
            min_temp = median(temp[temp == min(temp)])
  )
```
```{r}
mean_se_soil <- soil_filtered %>% 
  group_by(ir_freq, w_amount) %>% 
  summarize(median = median(temp),
            se_median = sd(temp)/sqrt(n()),
            median_max = median(temp[temp == max(temp)]),
            se_max = sd(temp)/sqrt(n()),
            median_min = median(temp[temp == min(temp)]),
            se_min = sd(temp)/sqrt(n()))
```

```{r}
# Create plot
p_soil_abs <- ggplot(df_daily_median, aes(x = date, color = factor(w_amount))) +
  geom_line(aes(y = median_temp), data = df_daily_median) +
  geom_line(aes(y = min_temp), data = df_daily_median) +
  geom_line(aes(y = max_temp), data = df_daily_median) +
  facet_wrap(~ ir_freq,, nrow = 3) +
  ggtitle("Daily median, maximum, and minimum soil temperature") +
  xlab("Date") +
  ylab("Temperature (°C)")  +
  theme_minimal()
p_soil_abs
```

#Cowplot
```{r}
# Subset the data for each level of ir_freq
df_median_freq1 <- filter(df_daily_median, ir_freq == 1)
df_median_freq2 <- filter(df_daily_median, ir_freq == 2)
df_median_freq3 <- filter(df_daily_median, ir_freq == 3)
```

```{r}
# Create a separate ggplot for each subset
plot_freq1 <- ggplot(df_median_freq1, aes(x = date, color = factor(w_amount))) +
  geom_line(aes(y = median_temp), data = df_median_freq1) +
  geom_line(aes(y = min_temp), data = df_median_freq1) +
  geom_line(aes(y = max_temp), data = df_median_freq1) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  xlab("Date") +
  ylab("Temperature (°C)")
```


```{r}
plot_freq2 <- ggplot(df_median_freq2, aes(x = date, color = factor(w_amount))) +
  geom_line(aes(y = median_temp), data = df_median_freq2) +
  geom_line(aes(y = min_temp), data = df_median_freq2) +
  geom_line(aes(y = max_temp), data = df_median_freq2) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  xlab("Date") +
  ylab("Temperature (°C)")
```


```{r}
plot_freq3 <- ggplot(df_median_freq3, aes(x = date, color = factor(w_amount))) +
  geom_line(aes(y = median_temp), data = df_median_freq3) +
  geom_line(aes(y = min_temp), data = df_median_freq3) +
  geom_line(aes(y = max_temp), data = df_median_freq3) +
  theme_minimal() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  xlab("Date") +
  ylab("Temperature (°C)")
```

#Read in irrigation data
```{r}
df_irrigation <- read_excel(paste0(path,"/data/raw/EasyManagement.xlsx"), sheet = "irr_minutes", col_names = TRUE, range = "A1:G37")


df_irrigation <- df_irrigation%>%
  slice(4:n()) %>%
  mutate(rel_irr2 = T2/T1,
         rel_irr3 = T3/T1
        )
```

#Make relative irrigaton plot
```{r}
plot_rel_irr <- ggplot(df_irrigation, aes(x = as.Date(Date))) +
  geom_line(aes(y = rel_irr2, color = "blue")) +
  geom_line(aes(y = rel_irr3, color = "red")) +
  scale_color_manual(values = c("blue" = "blue", "red" = "red"),
                     labels = c("Freq 2", "Freq 3")) +
  labs(color = "Frequency") +
  theme(axis.title.x = element_blank()) +
  ylab("Irrigation relative to frequency 1") +
  theme_minimal() +
  guides(color = FALSE)

plot_rel_irr
grob_rel_irr <- ggplotGrob(plot_rel_irr)
```




```{r, fig.width=10, fig.height=10}
library(cowplot)
#plot_grid(grob_rel_irr, grob_daily_temp, ncol = 1, height = c(0.1, 4))
# create a theme function to modify the plot margin 
theme_no_xaxis <- function(){
 #  theme_classic()+
  theme(axis.title.x = element_blank(), # remove unnecessary x axis
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(),
        legend.position = "none", # strip the legend
        # set plot margin to 0 
        plot.margin = unit(c(0,0,0,0), "mm"))
}

# a new irrigation plot for the composite 
plot_rel_irr2 <- ggplot(df_irrigation, aes(x = as.Date(Date))) +
  geom_line(aes(y = rel_irr2, color = "one")) +
  geom_line(aes(y = rel_irr3, color = "two")) +
  scale_color_manual(values = c(one = "blue",  two = "red"),
                     labels = c("Freq 2", "Freq 3")) +
  scale_x_date(limits = range(df_daily_median$date))+
  labs(color = "Frequency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) +
  theme_no_xaxis() +
  # theme(plot.margin = unit(c(1,1,1,1), "lines"))+
  ylab("Irrigation relative \n to frequency 1")
# get the legend from the first irrigation plot
legend_irr <- get_legend(plot_rel_irr)
# get the legend for the frequence 
# legend_freq <- get_legend(plot_freq1)
# create the main plot 
main <- plot_grid(plot_rel_irr2, 
          NULL, # use null and negative values in the rel_height to reduce the gap
          p_soil_abs + 
            theme(legend.position = "none")+ 
            ggtitle("") +
            scale_color_manual(values = c("blue","red")),
          align = "v", ncol = 1, axis = "tb",
          labels = c("a", "", "b"), 
          label_x = 0.1, label_y = 0.9,
          rel_heights = c(1, -0.1, 3))
# assemble the legend
legend <- plot_grid(legend_irr, legend_freq, ncol = 1)
# assemble legend and main plot 
plot_grid(main, rel_widths = c(0.9, 0.1))
```


#Kruskal wallis
```{r}
kruskal.test(median_temp ~ interaction(w_amount, ir_freq), data = df_daily_median)
kruskal.test(max_temp ~ interaction(w_amount, ir_freq), data = df_daily_median)
kruskal.test(min_temp ~ interaction(w_amount, ir_freq), data = df_daily_median)
```

```{r}
pairwise.wilcox.test(df_daily_median$max_temp, df_daily_median$treatment, p.adjust.method = "bonferroni")
```

```{r}
# First, run the pairwise Wilcoxon rank sum test and store the results in an object
pairwise_test <- pairwise.wilcox.test(df_daily_median$max_temp, df_daily_median$treatment, p.adjust.method = "bonferroni")

# Load the multcompView package
library(multcompView)

# Use the multcompLetters function to get the letters for pairwise comparisons
letters <- multcompLetters(pairwise_test$p.value)
```


```{r}
# Create a box plot
ggplot(df_daily_median, aes(x = interaction(w_amount, ir_freq), y = max_temp)) +
  geom_boxplot() +
  xlab("Interaction of w_amount and ir_freq") +
  ylab("Max Temperature in soil") +
  ggtitle("Distribution of Max Temperature in soil by Group")

# Create a histogram
ggplot(df_daily_median, aes(x = max_temp, fill = interaction(w_amount, ir_freq))) +
  geom_histogram(bins = 30, alpha = 0.5) +
  xlab("Max Temperature in soil") +
  ylab("Frequency") +
  ggtitle("Distribution of Max Temperature in soil by Group") +
  facet_wrap(~interaction(w_amount, ir_freq))
```



#Boxplot of temperature over time facet wrapped by trt (time series of boxplots to see trendline and variance)

```{r}
soil_boxplot <- soil_filtered %>%
    mutate(date = as.Date(time)) %>%
    group_by(date)
ggplot(soil_boxplot, aes(x = factor(date), y = temp, color = treatment)) +
  geom_boxplot() +
  facet_wrap(~ treatment, nrow = 6)
```
```{r}
soil_boxplot <- soil_filtered %>%
    mutate(date = as.Date(time)) %>%
    group_by(date)
ggplot(soil_boxplot, aes(x = factor(date), y = temp, color = w_amount)) +
  geom_boxplot() +
  facet_wrap(~ ir_freq, nrow = 3)
```
```{r}
soil_boxplot <- soil_filtered %>%
    mutate(date = as.Date(time)) %>%
    group_by(date)
ggplot(soil_boxplot, aes(x = factor(date), y = temp, color = ir_freq)) +
  geom_boxplot() +
  facet_wrap(~ w_amount, nrow = 2)
```
#Resample data for hourly mean of the four replicates
```{r}
# Group data by three-hour intervals and calculate mean value
soil_filt_1h <- soil_filtered %>%
  group_by(treatment, w_amount, ir_freq, interval = cut(time, breaks = "1 hours")) %>%
  summarize(mean_temp = mean(temp, na.rm = TRUE))

soil_filt_1h$interval <- as.POSIXct(soil_filt_1h$interval, format = "%Y-%m-%d %H:%M:%S")

# View new data table
soil_filt_1h
```

#Plot hourly mean temperature over time, facet wrapped by irrigation frequency (100% and 65% in one plot)

```{r}
ggplot(soil_filt_1h, aes(x = interval, y = mean_temp, color = w_amount)) +
  geom_line() +
  geom_point(data = ir_dates, aes(x = freq_2, y = point2), color = "red") +
  geom_point(data = ir_dates, aes(x = freq_3, y = point3), color = "blue") +
  facet_wrap(~ ir_freq, nrow = 3) +
  ggtitle("Soil temperature over time (hourly)") +
  xlab("Date") +
  ylab("Temperature")
```

#Plot hourly mean temperature over time, facet wrapped by water amount (all frequencies in one plot)

```{r}
ggplot(soil_filt_1h, aes(x = interval, y = mean_temp, color = ir_freq)) +
  geom_line() +
  geom_point(data = ir_dates, aes(x = freq_2, y = point2), color = "red") +
  geom_point(data = ir_dates, aes(x = freq_3, y = point3), color = "blue") +
  facet_wrap(~ w_amount, nrow = 2) +
  ggtitle("Soil temperature over time (hourly)") +
  xlab("Date") +
  ylab("Temperature")
```
#Filter data to only look at one irrigation interval (+ one week to see "lag" of temperature changes)
```{r}
start_date1 <- as.POSIXct("2022-12-08 00:00:00")
end_date1 <- as.POSIXct("2023-01-03 00:00:00")
soil_1st_int <- subset(soil_filt_1h, interval > start_date1 & interval < end_date1)

start_date2 <- as.POSIXct("2022-12-28 00:00:00")
end_date2 <- as.POSIXct("2023-01-19 00:00:00")
soil_2nd_int <- subset(soil_filt_1h, interval > start_date2 & interval < end_date2)

start_date3 <- as.POSIXct("2023-01-13 00:00:00")
end_date3 <- as.POSIXct("2023-02-09 00:00:00")
soil_3rd_int <- subset(soil_filt_1h, interval > start_date3 & interval < end_date3)

start_date4 <- as.POSIXct("2023-02-02 00:00:00")
end_date4 <- as.POSIXct("2023-02-24 00:00:00")
soil_4th_int <- subset(soil_filt_1h, interval > start_date4 & interval < end_date4)
```

```{r}
ggplot(soil_2nd_int, aes(x = interval, y = mean_temp, color = ir_freq)) +
  geom_line() +
  #geom_vline(xintercept = as.numeric(as.POSIXct("2023-01-21 12:00:00"))) +
  #geom_point(x = c("2023-01-21 12:00:00", "2023-01-27 12:00:00"), y = c(20, 21))
  geom_point(aes(x = as.POSIXct("2023-01-01 12:00:00"), y = 20), color = "red") +
  geom_point(aes(x = as.POSIXct("2023-01-07 12:00:00"), y = 20), color = "blue") +
  facet_wrap(~ w_amount, nrow = 2) +
  ggtitle("Soil temperature over time (hourly, 2nd irrigation interval)") +
  xlab("Date") +
  ylab("Temperature")
```
```{r}
ggplot(soil_2nd_int, aes(x = interval, y = mean_temp, color = w_amount)) +
  geom_line() +
  #geom_vline(xintercept = as.numeric(as.POSIXct("2023-01-21 12:00:00"))) +
  #geom_point(x = c("2023-01-21 12:00:00", "2023-01-27 12:00:00"), y = c(20, 21))
  geom_point(aes(x = as.POSIXct("2023-01-01 12:00:00"), y = 20), color = "red") +
  geom_point(aes(x = as.POSIXct("2023-01-07 12:00:00"), y = 20), color = "blue") +
  facet_wrap(~ ir_freq, nrow = 3) +
  ggtitle("Soil temperature over time (hourly, 2nd irrigation interval)") +
  xlab("Date") +
  ylab("Temperature")
```

```{r}
ggplot(soil_3rd_int, aes(x = interval, y = mean_temp, color = ir_freq)) +
  geom_line() +
  #geom_vline(xintercept = as.numeric(as.POSIXct("2023-01-21 12:00:00"))) +
  #geom_point(x = c("2023-01-21 12:00:00", "2023-01-27 12:00:00"), y = c(20, 21))
  geom_point(aes(x = as.POSIXct("2023-01-21 12:00:00"), y = 20), color = "red") +
  geom_point(aes(x = as.POSIXct("2023-01-27 12:00:00"), y = 20), color = "blue") +
  facet_wrap(~ w_amount, nrow = 2) +
  ggtitle("Soil temperature over time (hourly, 3rd irrigation interval)") +
  xlab("Date") +
  ylab("Temperature")
```
```{r}
ggplot(soil_3rd_int, aes(x = interval, y = mean_temp, color = w_amount)) +
  geom_line() +
  #geom_vline(xintercept = as.numeric(as.POSIXct("2023-01-21 12:00:00"))) +
  #geom_point(x = c("2023-01-21 12:00:00", "2023-01-27 12:00:00"), y = c(20, 21))
  geom_point(aes(x = as.POSIXct("2023-01-21 12:00:00"), y = 20), color = "red") +
  geom_point(aes(x = as.POSIXct("2023-01-27 12:00:00"), y = 20), color = "blue") +
  facet_wrap(~ ir_freq, nrow = 3) +
  ggtitle("Soil temperature over time (hourly, 3rd irrigation interval)") +
  xlab("Date") +
  ylab("Temperature")
```

#Resample data to larger intervalls to reduce business
```{r}
# Group data by day and summarize values
daily_data <- soil_filtered %>%
  mutate(date = as.Date(time)) %>%
  group_by(date, treatment) %>%
  summarize(temp = mean(temp))

# View new data table
daily_data
```

#Resample data to reduce business (3-hourly, currently loses most other information)
```{r}
# Group data by three-hour intervals and calculate mean value
soil_filt_3h <- soil_filtered %>%
  group_by(treatment, w_amount, ir_freq, interval = cut(time, breaks = "3 hours")) %>%
  summarize(mean_temp = mean(temp, na.rm = TRUE))

soil_filt_3h$interval <- as.POSIXct(soil_filt_3h$interval, format = "%Y-%m-%d %H:%M:%S")

# View new data table
soil_filt_3h
```
```{r}
ggplot(soil_filt_3h, aes(x = interval, y = mean_temp, color = treatment)) +
  geom_point() +
  facet_wrap(~ treatment, nrow = 6)
```
```{r}
ggplot(soil_filt_3h, aes(x = interval, y = mean_temp, color = treatment)) +
  geom_line()
```

```{r}
ggplot(soil_filtered, aes(x = time, y = temp, color = w_amount)) +
  geom_line() +
  geom_point(data = ir_dates, aes(x = freq_2, y = point2), color = "red") +
  geom_point(data = ir_dates, aes(x = freq_3, y = point3), color = "blue") +
  facet_wrap(~ ir_freq, nrow = 3) +
  ggtitle("Soil temperature over time (hourly)") +
  xlab("Date") +
  ylab("Temperature")
```

```{r}
ggplot(soil_filtered, aes(x = time, y = temp, color = ir_freq)) +
  geom_line() +
  geom_point(data = ir_dates, aes(x = freq_2, y = point2), color = "red") +
  geom_point(data = ir_dates, aes(x = freq_3, y = point3), color = "blue") +
  facet_wrap(~ w_amount, nrow = 2) +
  ggtitle("Soil temperature over time (hourly)") +
  xlab("Date") +
  ylab("Temperature")
```

```{r}
ggplot(soil_filt_3h, aes(x = interval, y = mean_temp, color = treatment)) +
  geom_line() +
  geom_vline(xintercept = as.numeric(as.POSIXct(c("2022-12-22 12:00:00", "2022-01-07 12:00:00", "2022-01-27 12:00:00", "2022-02-10 12:00:00")))) +
  facet_wrap(~ treatment)
```

#Plot 3-hourly data separated by amount and frequency
```{r}
ggplot(soil_filt_3h, aes(x = interval, y = mean_temp, color = w_amount)) +
  geom_line() +
  geom_point(data = ir_dates, aes(x = freq_2, y = point2), color = "red") +
  geom_point(data = ir_dates, aes(x = freq_3, y = point3), color = "blue") +
  facet_wrap(~ ir_freq, nrow = 3) +
  ggtitle("Soil temperature over time (3-hourly)") +
  xlab("Date") +
  ylab("Temperature")
```

```{r}
ggplot(soil_filt_3h, aes(x = interval, y = mean_temp, color = ir_freq)) +
  geom_line() +
  geom_point(data = ir_dates, aes(x = freq_2, y = point2), color = "red") +
  geom_point(data = ir_dates, aes(x = freq_3, y = point3), color = "blue") +
  facet_wrap(~ w_amount, nrow = 2) +
  ggtitle("Soil temperature over time (3-hourly)") +
  xlab("Date") +
  ylab("Temperature")

```

#Resample data to reduce business (6-hourly, currently loses most other information)
```{r}
# Group data by three-hour intervals and calculate mean value
soil_filt_6h <- soil_filtered %>%
  group_by(treatment, w_amount, ir_freq, interval = cut(time, breaks = "6 hours")) %>%
  summarize(mean_temp = mean(temp, na.rm = TRUE))

soil_filt_6h$interval <- as.POSIXct(soil_filt_6h$interval, format = "%Y-%m-%d %H:%M:%S")

# View new data table
soil_filt_6h
```

```{r}
ggplot(soil_filt_6h, aes(x = interval, y = mean_temp, color = ir_freq)) +
  geom_line() +
  facet_wrap(~ w_amount, nrow = 2) +
  ggtitle("Soil temperature over time (6-hourly)") +
  xlab("Date") +
  ylab("Temperature") +
  geom_vline(xintercept = c("2022-12-22 12:00:00", "2022-01-07 12:00:00", "2022-01-27 12:00:00", "2022-02-10 12:00:00"), 
             color = "red", linetype = "dashed")
```

```{r}
ggplot(soil_filt_6h, aes(x = interval, y = mean_temp, color = w_amount)) +
  geom_line() +
  facet_wrap(~ ir_freq, nrow = 3) +
  ggtitle("Soil temperature over time (6-hourly)") +
  xlab("Date") +
  ylab("Temperature")
```
