---
title: "Jochum_JAE2017_Reproducable"
author: "Tom MOore"
date:  '`r Sys.Date()`'
output: html_notebook
---

Install packages and functions
```{r setpu, include = FALSE}
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)}

packages <- c( "tidyverse","readr","here")
ipak(packages)
```


# Overview 

The analysis was carried out in R version `r version$version.string`. 

The analysis from the paper is reproduced here.

# Datasets

From https://datadryad.org/stash/dataset/doi:10.5061/dryad.qn119 downloaded the following files:

+ Jochum_JAE2017_Fig1_richness_coefs_confint.csv
+ Jochum_JAE2017_consumer_biomass.csv
+ Jochum_JAE2017_observed_consumer_richness.csv
+ Jochum_JAE2017_Fig2_biomass_coefs_confint.csv
+ Jochum_JAE2017_extrapolated_consumer_richness.csv
+ Jochum_JAE2017_predictors.csv

```{r}
fig1.coefs <- read.csv(file=here::here("Data/Jochum_2017/Jochum_JAE2017_Fig1_richness_coefs_confint.csv"))
```

# Simulate input data
```{r}
set.seed(1234)

# Observations
x <- 10

df <- tibble::tibble(
      LM = rnorm(x, mean = 10, sd = 4), # Adjusted to make the output more ecologically meaningful
   prich = rpois(x,5),  
      pH = rpois(x,5), # coefficient = 0 
  'C:N'  = rnorm(x),
  'C:P'  = rnorm(x, sd = 5), # Adjusted to make the output more ecologically meaningful
  'C:K'  = rnorm(x),
  'C:Ca' = rnorm(x),
  'C:Mg' = rnorm(x),
  'C:Na' = rnorm(x),
  'C:S'  = rnorm(x)
   )

write.csv(df, file =here::here("Data/df_input_data.csv"))
```


# Species richness calculation

+ Coefficients extracted from global model not for individual taxonomic groups
+ Coefficients were based on standardized (x-mean/SD) predictors so output has no ecological meaning.
+ Most influential coefficients are LM & C:P 
+ No intercept provided

+ To get more "realistic" coefficients - would need to rerun analysis based on non-standardized data and back transform (exp).  
```{r}
Spp.Rich.coeff <- fig1.coefs %>% 
                  filter(extrap_richness_response_group == 'all') %>%
                  select(2:3)


  df %>% 
  # select the correct predictor
  select(Spp.Rich.coeff$predictor) %>% 
  # give each combination an unique id
  mutate(row_id = 1:n()) %>% 
  # transform to long format for calculation
  pivot_longer(!row_id, names_to = "predictor") %>% 
  # join the coefs
  right_join(Spp.Rich.coeff) %>% 
  # get the product of value and its coef
  mutate(prod = value * coef) %>% 
  # group by the combination of predicators
  group_by(row_id)  %>% 
  # get the product for each combination and round to integer
  summarise(richness = round(sum(prod), digits = 0)) %>%
  # remove negative values
  mutate(richness = if_else(richness < 0, 0, richness),
         row_id = paste0("Site ", row_id ))
  
  
  
```
